---
title: "hydrological_data"
format: html
editor: visual
---

## Hydrological Data

```{r}
install.packages("dataRetrieval")
install.packages("purrr")
library(dataRetrieval)
library(ggplot2)
library(tidyverse)

```

```{r}
#Ventura River

siteNumber <- "11118500"
parameterCd <- "00060"
startDate <- "2023-08-19"
endDate <- "2023-08-22"

discharge <- readNWISdv(siteNumber, parameterCd, startDate, endDate)
```

-   Plot the time-series usgin `ggplot2`

```{r}

ggplot(data = discharge, aes(x = Date, y = X_00060_00003)) +
  geom_line()
```

-   What was the maximum discharge during this period of time?

August 21st

How would you try to determine when this stream gauge record started - (read the documentation :) ) ?

-   examine documentation, function in package askign for attirbutes of gauge

    ```{r}
    #siteNumbers <- c("01491000", "01645000")
    #siteINFO <- readNWISsite(siteNumbers)
    ```

-   How does the max value compare to the discharge during the storm on 2023-01-09 ?

```{r}

siteNumber <- "11118500"
parameterCd <- "00060"
startDate <- "2023-08-19"
endDate <- "2023-08-22"

discharge <- readNWISdv(siteNumber, parameterCd, startDate, endDate)


#replace dates for the storm 
ggplot(data = discharge, aes(x = Date, y = X_00060_00003)) +
  geom_line()


```

\
Example 2: Meta jam

Let's determine what percentage of Alaskan household are speaking only English!

The data: https://doi.org/10.5063/F1N58JPP

```{r}
install.packages("metajam")
library(metajam)

```

Everytime you run data you re-request code!

```{r}

# set inputs
data_obj <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
path <- "~/R"
# download data and metadata
download_d1_data(data_obj, path)

```

```{r}
# R base can do this
download.file(data_obj, "Alaska_Schools_Rentention2009_15.csv")
# or even 
my_data <- read.csv(data_obj)



```

```{r}
my_data <- 
```

```{r}
url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"

#to save it here use . 
local_path <- "." 

metajam::download_d1_data(url, local_path)


```

\^ this command took the full_metadata file and placed it into constructive tables

Read csv into R

```{r}


ak_household_data <- metajam::read_d1_files("doi_10.5063_F1N58JPP__household_language__csv/")

```

```{r}
data <- ak_household_data$data
View(data)
```

-   compute the percentage of Alaskan household speaking only English from 2009 to 2015

1.  Create a plot to visualize this data

```{r}
data_percent <- data %>%
  mutate(english_only_percent = speak_only_english/total) %>%
  filter(Year %in% c(2009:2015)) %>%
  group_by(Year) %>%
  summarise(english_only_percent = mean(english_only_percent, na.rm = TRUE))

```

```{r}
ggplot(data = data_percent, aes(x = Year, y = english_only_percent )) +
  geom_line()
```
